Improving the templates for multicolor Ramsey numbers and Schur numbers

### 2022-06-26 article_ramsey/template_improvement.tex
definitions

\section{Introduction}
This article addresses the edge-coloring of complete graphs in an arbitrary number of colors while limiting the clique 
number in each color, as well as sum-free colorings of positive integers. A pattern based construction which uses a 
particular type of coloring nammed sf-template was described in \cite{RowleyRamsey}. This construction was used to 
improve lower bounds on several Ramsey numbers as well as Schur numbers. Lower bounds were then improved by describing 
larger templates in \cite{schurboyz} for Schur numbers and in \cite{rowleyramseysat} for multicolor Ramsey numbers.

@@

Theorem \ref{thm:linkintegers} shows a link between Schur numbers and Ramsey numbers:
\[\forall n \in \mathbb{N}^*, S(n) = L_n(3) \leqslant R_n(3) - 2\].

\section{\(b\)-Templates}
\label{sec:temp}
Dans cette section, on introduit les \(b\)-templates. Les templates présentés dans la section \ref{sec:b-temp} vont apparaître comme un cas particulier des \(b\)-templates puisqu'ils correspondent au cas \(b = 0\).

On commence par décrire \(\TS_{b,k}\), un ensemble particulier de \(k\)-tuples qui sera utilisé dans la définition des \(b\)-templates.

\begin{definition}[Arbre associé à un tuple]
Soit \(a \in \mathbb{N}^*\), soit \(b \in \mathbb{N}\). A chaque tuple d'entiers, on associe un arbre de manière récurssive :
\[ \left\{
\begin{array}{l}
	\forall x \in \mathbb{R}, \tree(x) = x, \\
	\forall k \in [\![2, +\infty[\![, \forall x \in \mathbb{R}^k,
		\left\{
		\begin{array}{l l}
			\tree(x) = (x_1, \tree(x_2, ..., x_k)) &  ~\text{si}~ x_2 - x_1 > b, \\
			\tree(x) = (x_1, \tree(x_2, ..., x_k), \tree(x_2 + a, ..., x_k)) &  ~\text{si}~ 0 \leqslant x_2 - x_1 \leqslant b, \\
			\tree(x) = \tree\left(x_1, x_2 + a \left\lceil \dfrac{x_1 - x_2}{a}\right\rceil, x_3, ..., x_k\right) & ~\text{si}~ x_2 < x_1.
		\end{array}
		\right.
\end{array}
\right. \]
Pour un arbre \(\tree(x)\), on note \(\treeset(x)\) l'ensemble des tuples d'entiers que l'on peut construire en collectant les labels au cours d'une descente de cet arbre.
\end{definition}

\begin{definition}[Ensembles \(S_{b,k}(A)\) et \(\TS_{b,k}(A)\)]
Soit \(A \subset \mathbb{N}^*\), soit \(b \in \mathbb{N}\) et soit \(k \in [\![2, +\infty[\![\). On note \(S_{b,k}(A)\) l'ensemble de \(k\)-tuples  à valeurs distinctes dans \(A\) et dont les éléments de \([\![1, b]\!]\) sont situés au début et apparaissent dans l'ordre croissant :\\
\(S_{b,k}(A) = \left\{ x \in \mathfrak{S}_k(A) ~:~ \forall i \in  [\![2, k]\!], x_i \leqslant b \implies x_i \geqslant x_{i -1} \right\}\). On définit alors un ensemble \(\TS_{b,k}(A)\) de \(k\)-tuples : \(\TS_{b,k}(A) = \displaystyle\bigcup\limits_{x \in S_{b,k}(A)} \treeset(x)\).
\end{definition}

On définit aussi un opérateur de projection de \(\mathbb{Z}\) vers \([\![1, a + b]\!]\).

\begin{definition}[Projection \(\pi_{a, b}\)]
Soit \(a \in \mathbb{N}^*\) et soit \(b \in \mathbb{N}\). On définit une projection, noté \(\pi_{a, b}\) ou plus simplement \(\pi\) si il n'y a pas d'ambiguïté, sur \([\![1, a + b]\!]\) de la manière suivante.
\[ \begin{array}{c c c l}
	\pi_{a, b} : & \mathbb{Z} & \longrightarrow & [\![1, a + b]\!] \\
	 & x & \longmapsto & 
		\left\{
		\begin{array}{l l}
			x & ~\text{si}~ 1 \leqslant x \leqslant b \\
			(x \mod a) + \mathds{1}_{[\![0, b]\!]}(x \mod a) & ~\text{si}~ x > b
		\end{array}
		\right.
\end{array} \]
\end{definition}

On peut maintenant définir les \(b\)-templates.

\begin{definition}[\(b\)-Templates]
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {[\![3, +\infty[\![}^r\). Soit \(a \in \mathbb{N}^*\) et soit \(b \in \mathbb{N}\).  Un \(b\)-template de largeur \(a\) à \(r + 1\) couleurs, est défini comme une partition de \([\![1, a + b]\!]\) en \(r+1\) sous-ensembles \(A_1, ..., A_{r+1}\) vérifiant :
\begin{enumerate}[(i)]
\item \(a \in A_{r+1}\),
\item \(\forall (x, y) \in A_{r + 1}^2, x + y \notin [\![a + b + 1, 2 a + b]\!] \implies \pi(x + y) \notin A_{r + 1}\),
\item \(\forall c \in [\![1, c]\!], \forall x \in \TS_{b, k_c - 1}(A_c), \exists (i, j) \in {[\![1, k_c - 1]\!]}^2, i < j ~\text{et}~ \pi(x_j - x_i) \notin A_c\).
\end{enumerate}
 \(T_b(k_1, ..., k_r, t; a)\) désigne l'un (quelconque) de ces \(b\)-templates et l'ensemble de ces \(b\)-templates est noté \(\mathcal{T}_b(k_1, ..., k_r, t; a)\).  La couleur \(r + 1\) joue un rôle particulier et elle est appelée "couleur template".
\end{definition}

La couleur \(r + 1\)  n'est pas nécessairement la dernière par ordre d'apparition, la désignation de \(r+1\) comme couleur template signalée par \(t\) est une convention qui allège les notations en évitant de devoir écrire \(\mathcal{T}_b(k_1, ..., k_{i - 1}, t,  k_{i + 1}, ..., k_r; n)\) par exemple.

Dans le cas où tous les \(k_c\) sont égaux à 3, ces \(b\)-templates constituent des \(b\)-S-templates applicables pour les nombres de Schur similaires au \(b\)-WS-templates pour les nombres de Schur faibles définis dans \cite{schurboyz}.

\begin{proposition}[Lien avec les templates]
Soit \(r \in \mathbb{N}\), soit \((k_c)_{1 \leqslant c \leqslant r} \in {[\![3, +\infty[\![}^r\) et soit \(a \in \mathbb{N}^*\). Alors :
\[ \mathcal{T}(k_1, ..., k_r, t; a) = \mathcal{T}_0(k_1, ..., k_r, t; a). \]
\end{proposition}

\begin{proposition}
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {[\![3, +\infty[\![}^r\). Soit \(a \in \mathbb{N}^*\) et soit \(b \in \mathbb{N}\). Alors  \(T_b(k_1, ..., k_r, t; a) \subset L(k_1, ..., k_r, 3; a + b)\).
\end{proposition}

\begin{definition}[Coloriage par défaut]
Soit \(n \in \mathbb{N}^*\), soit \(b \in \mathbb{N}\) et soit \(r \in \mathbb{N}\). Soit \(H\) un coloriage de \([\![1, n]\!]\) à \(r +1\) couleurs. On pose \(d = \min \left(H^{-1}(\{r + 1\}) \backslash [\![1, b]\!] \right) - b  - 1\). On définit alors un coloriage \(\coldef_b(H)\), dit par défaut, associé au coloriage \(H\) :
\[\begin{array}{c c c c}
	\coldef_b(H) : & [\![1, d]\!] & \longrightarrow & [\![1, r + 1]\!] \\
	 & x & \longmapsto & H(x + b)
\end{array}\]
\end{definition}

\begin{definition}[Méthode de construction]
Soit \((a, p) \in {\mathbb{N}^*}^2\) et soit \(b \in \mathbb{N}\). Soit \((r, s) \in {\mathbb{N}^*}^2\). On se donne un coloriage de \([\![1, a + b]\!]\) à \(r + 1\) couleurs (appelée \(H\), pour horizontal), et un coloriage de \([\![1, p]\!]\) à \(s\) couleurs (appelée \(V\), pour vertical).

Pour un \(d \in \mathbb{N}\), on se donne aussi un coloriage dit final \(F\) de \([\![1, d]\!]\) à \(r + 1\) couleurs. Il est parfois utile de définir \(F\) au cas par cas, mais on peut aussi lui attribuer une valeur par défaut \(F = \coldef_b(H)\). Dans le cas où \(F\) n'est pas utilisé, on note \(F = \varnothing\) l'application vide.

On construit alors un coloriage \(\gamma\) de \([\![1, a \times p + b + d]\!]\) à \(r + s\) couleurs comme suit. De manière informelle, on commence par utiliser le coloriage initial \(I \defeq H_{| \, [\![1, b]\!]}\), puis on répéte \(p\) fois le pattern \(P \defeq H_{| \, [\![b + 1, a + b]\!]}\) en remplaçant pour la \(i\)-ème itération la couleur spéciale par \(r + \mathcal{V}(i)\), et finalement on ajoute le motif formé l'enchaînement des couleurs dans \(\mathcal{F}\). Formellement :
\[\begin{array}{c c c l}
	\gamma: & [\![1, a \times p + b + d]\!]  & \longrightarrow &  [\![1, r + s]\!] \\
 	& x & \longmapsto & 
		\left\{\begin{array}{l l}
			H(x) & ~\text{si}~ x \leqslant b \\
			H(\pi(x)) & ~\text{si}~  b + 1 \leqslant x \leqslant a \times p + b ~\text{et}~ H(\pi(x)) \neq r + 1 \\
			r + V \left(\left\lceil \dfrac{x - b}{a} \right\rceil\right) & ~\text{si}~ b + 1 \leqslant x \leqslant a \times p + b ~\text{et}~ H(\pi(x)) = r + 1 \\
			F(x - a \times p - b) & ~\text{si}~ x \geqslant a \times p + b + 1
		\end{array}\right.
\end{array}\]
\end{definition}

\begin{figure}[H]
\caption{Exemple de construction avec \(b = 5\)}
\renewcommand{\arraystretch}{1.7}
\begin{center}

\begin{tabular}{c}
	\begin{NiceTabular}{c}[hvlines]
		\CodeBefore
			\cellcolor{cyan}{1-1}
		\Body
			couleur template
	\end{NiceTabular} \\
	coloriage \(H\) \\
	\(
		\begin{NiceArray}{*{17}{c}}[hvlines,last-row]
		\CodeBefore
			\cellcolor{red}{1-1}
			\cellcolor{red}{1-2}
			\cellcolor{cyan}{1-3}
			\cellcolor{red}{1-4}
			\cellcolor{green}{1-5}
			\cellcolor{green}{1-6}
			\cellcolor{red}{1-7}
			\cellcolor{green}{1-8}
			\cellcolor{cyan}{1-9}
			\cellcolor{cyan}{1-10}
			\cellcolor{cyan}{1-11}
			\cellcolor{cyan}{1-12}
			\cellcolor{cyan}{1-13}
			\cellcolor{red}{1-14}
			\cellcolor{green}{1-15}
			\cellcolor{cyan}{1-16}
			\cellcolor{red}{1-17}
		\Body
			\,\,1\,\, & \,\,2\,\, & \,\,3\,\, & \,\,4\,\, & \,\,5\,\, & \,\,6\,\, & \,\,7\,\, & \,\,8\,\, & \,\,9\,\, & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 \\
			\Block{1-5}{\underbrace{\hphantom{................................}}_{\textstyle I}} & & & & & \Block{1-12}{\underbrace{\hphantom{..................................................................................}}_{\textstyle P}}
		\end{NiceArray}\) \\
	\begin{tabular}{c c}
		coloriage \(V\) & coloriage \(F\) \\
		\(
			\begin{NiceArray}{c}[hvlines]
			\CodeBefore
				\cellcolor{cyan}{1-1}
				\cellcolor{yellow}{2-1}
				\cellcolor{yellow}{3-1}
				\cellcolor{cyan}{4-1}
			\Body
				\,\,1\,\, \\
				\,\,2\,\, \\
				\,\,3\,\, \\
				\,\,4\,\, \\
			\end{NiceArray}\) &
			\( \begin{NiceArray}{*{4}{c}}[hvlines]
			\CodeBefore
				\cellcolor{red}{1-1}
				\cellcolor{cyan}{1-2}
				\cellcolor{green}{1-3}
				\cellcolor{green}{1-4}
			\Body
				\,\,1\,\, & \,\,2\,\, & \,\,3\,\, & \,\,4\,\,
			\end{NiceArray}\) \\	
	\end{tabular} \\
	\(\cons_5(H, V, F)\) \\
	\(
		\begin{NiceArray}{*{12}{c}}[corners={NW,SE},hvlines]
		\CodeBefore
		    	\cellcolor{red}{1-8}
			\cellcolor{red}{1-9}
			\cellcolor{cyan}{1-10}
			\cellcolor{red}{1-11}
			\cellcolor{green}{1-12}
			\cellcolor{green}{2-1}
			\cellcolor{red}{2-2}
			\cellcolor{green}{2-3}
			\cellcolor{cyan}{2-4}
			\cellcolor{cyan}{2-5}
			\cellcolor{cyan}{2-6}
			\cellcolor{cyan}{2-7}
			\cellcolor{cyan}{2-8}
			\cellcolor{red}{2-9}
			\cellcolor{green}{2-10}
			\cellcolor{cyan}{2-11}
			\cellcolor{red}{2-12}
			\cellcolor{green}{3-1}
			\cellcolor{red}{3-2}
			\cellcolor{green}{3-3}
			\cellcolor{yellow}{3-4}
			\cellcolor{yellow}{3-5}
			\cellcolor{yellow}{3-6}
			\cellcolor{yellow}{3-7}
			\cellcolor{yellow}{3-8}
			\cellcolor{red}{3-9}
			\cellcolor{green}{3-10}
			\cellcolor{yellow}{3-11}
			\cellcolor{red}{3-12}
			\cellcolor{green}{4-1}
			\cellcolor{red}{4-2}
			\cellcolor{green}{4-3}
			\cellcolor{yellow}{4-4}
			\cellcolor{yellow}{4-5}
			\cellcolor{yellow}{4-6}
			\cellcolor{yellow}{4-7}
			\cellcolor{yellow}{4-8}
			\cellcolor{red}{4-9}
			\cellcolor{green}{4-10}
			\cellcolor{yellow}{4-11}
			\cellcolor{red}{4-12}
			\cellcolor{green}{5-1}
			\cellcolor{red}{5-2}
			\cellcolor{green}{5-3}
			\cellcolor{cyan}{5-4}
			\cellcolor{cyan}{5-5}
			\cellcolor{cyan}{5-6}
			\cellcolor{cyan}{5-7}
			\cellcolor{cyan}{5-8}
			\cellcolor{red}{5-9}
			\cellcolor{green}{5-10}
			\cellcolor{cyan}{5-11}
			\cellcolor{red}{5-12}
			\cellcolor{red}{6-1}
			\cellcolor{cyan}{6-2}
			\cellcolor{green}{6-3}
			\cellcolor{green}{6-4}
		 \Body
		    	   & & & & & & & 1 &  2 & 3 &  4 &  5 \\
		    	 6 &  7 &  8 &  9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 \\
		    	18 & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 \\
		    	30 & 31 & 32 & 33 & 34 & 35 & 36 & 37 & 38 & 39 & 40 & 41 \\
			42 & 43 & 44 & 45 & 46 & 47 & 48 & 49 & 50 & 51 & 51 & 53 \\
			54 & 55 & 56 & 57 \\
		\end{NiceArray}\)
\end{tabular}
\end{center}
\end{figure}

\begin{definition}[Compatibilité avec un \(b\)-template]
Soit \(a \in \mathbb{N}^*\), soit \(b \in \mathbb{N}\) et soit \(r \in \mathbb{N}\). Soit \((k_c)_{1 \leqslant c \leqslant r} \in {[\![3, + \infty [\![}^r\). Soit \(T \in \mathcal{T}_b(k_1, ..., k_r; a)\) un \(b\)-template et soit \(A_1, ..., A_{r+1}\) la partition associée. Soit \(d \in [\![1, a]\!]\). Soit \(F\) un coloriage de \([\![1, d]\!]\) à \(r +1\) couleurs et soit \(B_1, ..., B_{r + 1}\) la partition associée. Le coloriage \(F\)  est dit compatible avec le coloriage \(T\) si il vérifie :
\begin{enumerate}[(i)]
\item \(\forall (x,y) \in A_{r+1}^2, (x + y > b ~\text{et}~ \pi(x + y) \leqslant d) \implies \pi(x + y)  \notin B_{r+1}\)
\item \(\forall (x,y) \in A_{r+1} \times B_{r+1}, x + y \leqslant d \implies x + y  \notin B_{r+1}\)
\item \(\forall c \in [\![1, r]\!], \forall k \in [\![1, k_c-1]\!], \\
	\left\{ \text{ou} 
	\begin{array}{l}
		\exists (i, j) \in {[\![1, k]\!]}^2, |x_i - x_j| \notin A_c, \\
		TODO \\
		\exists (i, j) \in {[\![1, k_c - k - 1]\!]}^2, i < j ~\text{et}~ \pi(y_j - y_i) \notin A_c.
	\end{array}
	\right.\)
\end{enumerate}
\end{definition}

>>>>

\section{Introduction}

This article addresses the edge-coloring of complete graphs in an arbitrary number of colors while limiting the clique 
number in each color, as well as sum-free colorings of positive integers. A pattern based construction which uses a 
particular type of coloring nammed sf-template was described in \cite{RowleyRamsey}. This construction was used to 
improve lower bounds on several Ramsey numbers as well as Schur numbers. Lower bounds were then improved by describing 
larger templates in \cite{schurboyz} for Schur numbers and in \cite{rowleyramseysat} for multicolor Ramsey numbers.

@@

Theorem \ref{thm:linkintegers} shows a link between Schur numbers and Ramsey numbers:
\[\forall n \in \mathbb{N}^*, S(n) = L_n(3) \leqslant R_n(3) - 2.\]

\section{\(b\)-Templates}
\label{sec:temp}

\begin{definition}[Construction method]
Let \((a, p) \in {\mathbb{N}^*}^2\) and let \(b \in \mathbb{N}\). Let \((r, s) \in {\mathbb{N}^*}^2\). Let \(H\) 
(standing for "horizontal") be a coloring of \([\![1, a + b]\!]\) with \(r + 1\) colors, and let \(V\)  (standing for 
"vertical") be a coloring of \([\![1, p]\!]\) with \(s\) colors. Let \(d \in \mathbb{N}\) and let \(F\) (standing for 
"final") be a coloring of \([\![1, d]\!]\) with \(r + 1\) colors. The empty coloring (i.e. \(d = 0\)) is denoted by 
\(F = \varnothing\). The construction produces a coloring \(\cons_b(H, V, F)\) of \([\![1, a p + b + d]\!]\) with 
\(r + s\) colors.

Informally, one starts by using the initial coloring \(I \defeq H_{| \, [\![1, b]\!]}\), then repeats \(p\) times the 
pattern \(P \defeq H_{| \, [\![b + 1, a + b]\!]}\) while replacing at the \(i\)\textsuperscript{th} iteration the color 
\(r + 1\) by \(r + \mathcal{V}(i)\), and eventually appends the coloring \(\mathcal{F}\). Formally:
\[\begin{array}{ccl}
	[\![1, a p + b + d]\!]  & \longrightarrow &  [\![1, r + s]\!] \\
 	x & \longmapsto & 
		\left\{\begin{array}{l l}
			H(x) & ~\text{if}~ x \leqslant b \\
			H(\pi(x)) & ~\text{if}~  b + 1 \leqslant x \leqslant a p + b ~\text{and}~ H(\pi(x)) \neq r + 1 \\
			r + V \left(\left\lceil \dfrac{x - b}{a} \right\rceil\right) & ~\text{if}~ b + 1 \leqslant x \leqslant a p + b 
				~\text{and}~ H(\pi(x)) = r + 1 \\
			F(x - a p - b) & ~\text{if}~ x \geqslant a p + b + 1
		\end{array}\right.
\end{array}\]
\end{definition}

\begin{figure}[H]
\caption{Example of construction with \(b = 5\)}
\renewcommand{\arraystretch}{1.6}
\begin{center}

\begin{tabular}{c}
	\begin{NiceTabular}{c}[hvlines]
		\CodeBefore
			\cellcolor{cyan}{1-1}
		\Body
			template color
	\end{NiceTabular} \\
	\(H\) coloring \\
	\(
		\begin{NiceArray}{*{17}{c}}[hvlines,last-row]
		\CodeBefore
			\cellcolor{red}{1-1}
			\cellcolor{red}{1-2}
			\cellcolor{cyan}{1-3}
			\cellcolor{red}{1-4}
			\cellcolor{green}{1-5}
			\cellcolor{green}{1-6}
			\cellcolor{red}{1-7}
			\cellcolor{green}{1-8}
			\cellcolor{cyan}{1-9}
			\cellcolor{cyan}{1-10}
			\cellcolor{cyan}{1-11}
			\cellcolor{cyan}{1-12}
			\cellcolor{cyan}{1-13}
			\cellcolor{red}{1-14}
			\cellcolor{green}{1-15}
			\cellcolor{cyan}{1-16}
			\cellcolor{red}{1-17}
		\Body
			\,\,1\,\, & \,\,2\,\, & \,\,3\,\, & \,\,4\,\, & \,\,5\,\, & \,\,6\,\, & \,\,7\,\, & \,\,8\,\, & \,\,9\,\, & 10 & 
				11 & 12 & 13 & 14 & 15 & 16 & 17 \\
			\Block{1-5}{\underbrace{\hphantom{................................}}_{\textstyle I}} & & & & & \Block{1-12}{\underbrace{\hphantom{..................................................................................}}_{\textstyle P}}
		\end{NiceArray}\) \\
	\begin{tabular}{c c}
		\begin{tabular}{c}
			\(V\) coloring \\
			\(
				\begin{NiceArray}{c}[hvlines]
				\CodeBefore
					\cellcolor{cyan}{1-1}
					\cellcolor{yellow}{2-1}
					\cellcolor{yellow}{3-1}
					\cellcolor{cyan}{4-1}
				\Body
					\,\,1\,\, \\
					\,\,2\,\, \\
					\,\,3\,\, \\
					\,\,4\,\, \\
				\end{NiceArray}\)
		\end{tabular} &
		\begin{tabular}{c}
			\(F\) coloring \\
			\(
				\begin{NiceArray}{*{4}{c}}[hvlines]
				\CodeBefore
					\cellcolor{red}{1-1}
					\cellcolor{cyan}{1-2}
					\cellcolor{green}{1-3}
					\cellcolor{green}{1-4}
				\Body
					\,\,1\,\, & \,\,2\,\, & \,\,3\,\, & \,\,4\,\,
				\end{NiceArray}\)
		\end{tabular}
	\end{tabular} \\
	\(\cons_5(H, V, F)\) \\
	\(
		\begin{NiceArray}{*{12}{c}}[corners={NW,SE},hvlines]
		\CodeBefore
		    	\cellcolor{red}{1-8}
			\cellcolor{red}{1-9}
			\cellcolor{cyan}{1-10}
			\cellcolor{red}{1-11}
			\cellcolor{green}{1-12}
			\cellcolor{green}{2-1}
			\cellcolor{red}{2-2}
			\cellcolor{green}{2-3}
			\cellcolor{cyan}{2-4}
			\cellcolor{cyan}{2-5}
			\cellcolor{cyan}{2-6}
			\cellcolor{cyan}{2-7}
			\cellcolor{cyan}{2-8}
			\cellcolor{red}{2-9}
			\cellcolor{green}{2-10}
			\cellcolor{cyan}{2-11}
			\cellcolor{red}{2-12}
			\cellcolor{green}{3-1}
			\cellcolor{red}{3-2}
			\cellcolor{green}{3-3}
			\cellcolor{yellow}{3-4}
			\cellcolor{yellow}{3-5}
			\cellcolor{yellow}{3-6}
			\cellcolor{yellow}{3-7}
			\cellcolor{yellow}{3-8}
			\cellcolor{red}{3-9}
			\cellcolor{green}{3-10}
			\cellcolor{yellow}{3-11}
			\cellcolor{red}{3-12}
			\cellcolor{green}{4-1}
			\cellcolor{red}{4-2}
			\cellcolor{green}{4-3}
			\cellcolor{yellow}{4-4}
			\cellcolor{yellow}{4-5}
			\cellcolor{yellow}{4-6}
			\cellcolor{yellow}{4-7}
			\cellcolor{yellow}{4-8}
			\cellcolor{red}{4-9}
			\cellcolor{green}{4-10}
			\cellcolor{yellow}{4-11}
			\cellcolor{red}{4-12}
			\cellcolor{green}{5-1}
			\cellcolor{red}{5-2}
			\cellcolor{green}{5-3}
			\cellcolor{cyan}{5-4}
			\cellcolor{cyan}{5-5}
			\cellcolor{cyan}{5-6}
			\cellcolor{cyan}{5-7}
			\cellcolor{cyan}{5-8}
			\cellcolor{red}{5-9}
			\cellcolor{green}{5-10}
			\cellcolor{cyan}{5-11}
			\cellcolor{red}{5-12}
			\cellcolor{red}{6-1}
			\cellcolor{cyan}{6-2}
			\cellcolor{green}{6-3}
			\cellcolor{green}{6-4}
		 \Body
		    	   & & & & & & & 1 &  2 & 3 &  4 &  5 \\
		    	 6 &  7 &  8 &  9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 \\
		    	18 & 19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 & 28 & 29 \\
		    	30 & 31 & 32 & 33 & 34 & 35 & 36 & 37 & 38 & 39 & 40 & 41 \\
			42 & 43 & 44 & 45 & 46 & 47 & 48 & 49 & 50 & 51 & 51 & 53 \\
			54 & 55 & 56 & 57 \\
		\end{NiceArray}\)
\end{tabular}
\end{center}
\end{figure}

In some cases, it may be useful to use a specific coloring for the final coloring. However, a default coloring 
associated to the horizontal coloring can be used instead.

\begin{definition}[Default coloring]
Let \(n \in \mathbb{N}^*\), \(b \in \mathbb{N}\) and \(r \in \mathbb{N}\). Let \(H\) be a coloring of \([\![1, n]\!]\) 
with \(r +1\) colors. Set \(d = \min \left(H^{-1}(\{r + 1\}) \backslash [\![1, b]\!] \right) - b  - 1\). The default 
coloring \(\coldef_b(H)\)  associated to \(H\) is defined as follows.
\[\begin{array}{c c c c}
	\coldef_b(H) : & [\![1, d]\!] & \longrightarrow & [\![1, r + 1]\!] \\
	 & x & \longmapsto & H(x + b)
\end{array}\]
\end{definition}

The construction described in \cite{RowleyRamsey} for sf-templates corresponds to the case \(b = 0\) and it uses the 
default coloring. Contrary to sf-templates, \(b\)-templates are directly defined as particular colorings for which the 
clique numbers remain unchanged in the construction. The rest of this section defines the \(b\)-templates as well as 
the notion of compatibility for the final coloring.

In order to describe the configurations in which a given tuple can induce a monochromatic clique when using the 
construction, a tree is associated to every tuple.

\begin{definition}
A tree node has two components: the first one is the label and the second one is the set of descendants of this node. A 
branch of a tree \(T\) is a tuple \((x_i)_{1 \leqslant i \leqslant k}\) such that there is a sequence of trees 
\((t_i)_{1 \leqslant i \leqslant k}\) satisfying: \(t_1 = T\) and for all \(i \in [\![1, k]\!]\) \(x_i\) is the label 
of \(t_i\) and for all \(i \in [\![1, k - 1]\!]\) \(t_{i + 1}\) is a descendant of \(t_i\).

Let \(a \in \mathbb{N}^*\) and \(b \in \mathbb{N}\). The \(\tree_{a, b}\) application is recurssively defined as 
follows.
\[ \left\{
\begin{array}{l}
	\forall x \in \mathbb{Z}, \tree_{a, b}(x) = (x, \varnothing), \\
	\forall k \in [\![2, +\infty[\![, \forall x \in \mathbb{Z}^k, \\
	\mspace{30mu}
		\left\{
		\begin{array}{l l}
			\tree_{a, b}(x) = (x_1, \{\tree_{a, b}(x_2, ..., x_k)\}) &  ~\text{if}~ x_2 - x_1 > b \\
			\tree_{a, b}(x) = (x_1, \{\tree_{a, b}(x_2, ..., x_k), \tree_{a, b}(x_2 + a, ..., x_k)\}) &  ~\text{if}~ 0 \leqslant 
				x_2 - x_1 \leqslant b \\
			\tree_{a, b}(x) = \tree_{a, b} \left(x_1, x_2 + a \left \lceil \dfrac{x_1 - x_2}{a} \right \rceil, x_3, ..., x_k 
				\right) & ~\text{if}~ x_2 < x_1
		\end{array}
		\right.
\end{array}
\right. \]
Let \(x\) be a non-empty tuple of integers. The set of all the branches of the tree \(\tree_{a, b}(x)\) is denoted by 
\(\treeset_{a, b}(x)\).
\end{definition}

Let \(A \subset \mathbb{N}^*\) and let \(k \in [\![2, +\infty]\!]\). When using the construction, not every tuple in 
\(A^k\) have a tree which describes cliques that appear in the construction.

\begin{definition}
Let \(A \subset \mathbb{N}^*\), \(b \in \mathbb{N}\) and \(k \in [\![2, +\infty[\![\). The set of \(k\)-tuples of 
\(A^k\) whose values are pariwise distinct is denoted by \(\mathfrak{S}_k(A)\). The set of \(k\)-tuples of 
\(\mathfrak{S}_k(A)\) such that the elements of a tuple that are in \([\![1, b]\!]\) appear in increasing order at 
begining of this tuple is denoted by \(S_{b,k}(A)\). That is 
\(S_{b,k}(A) = \left\{ x \in \mathfrak{S}_k(A) ~:~ \forall i \in  [\![2, k]\!], x_i \leqslant b \implies x_i \geqslant 
x_{i -1} \right\}\). The set of branches of trees of tuples in \(S_{b, k}(A)\) is denoted by \(\TS_{b,k}(A)\): 
\(\TS_{b,k}(A) = \displaystyle \bigcup \limits_{x \in S_{b,k}(A)} \treeset(x)\).
\end{definition}

Most of the integers which appear in the tuples of \(\TS_{b,k}(A)\) (as well as the differences of these integers) are 
not in \([\![1, a + b]\!]\). The role of the projection \(\pi_{a, b}\) is to project these integers onto 
\([\![1, a + b]\!]\) in order to describe the constraints defning the \(b\)-templates.

\begin{definition}
Let \(a \in \mathbb{N}^*\) and \(b \in \mathbb{N}\).The projection \(\pi_{a, b}\) is defined as follows.
\[ \begin{array}{c c c l}
	\pi_{a, b} : & \mathbb{Z} & \longrightarrow & [\![1, a + b]\!] \\
	 & x & \longmapsto & 
		\left\{
		\begin{array}{l l}
			x & ~\text{if}~ 1 \leqslant x \leqslant b \\
			(x \mod a) + a \mathds{1}_{[\![0, b]\!]}(x \mod a) & ~\text{if}~ x > b
		\end{array}
		\right.
\end{array} \]
\end{definition}

The \(b\)-templates are designed to be used as the horizontal coloring in the construction.

\begin{definition}[\(b\)-Templates]
Let \(r \in \mathbb{N}\) and \((k_c)_{1 \leqslant c \leqslant r} \in {[\![3, +\infty[\![}^r\). Let 
\(a \in \mathbb{N}^*\) and \(b \in [\![0, a - 1]\!]\).  A \(b\)-template with width \(a\) and \(r + 1\) colors is 
defined as a partition of \([\![1, a + b]\!]\) into \(r+1\) subsets \(A_1, ..., A_{r+1}\) such that:
\begin{enumerate}[(i)]
\item \(a \in A_{r+1}\),
\item \(\forall (x, y) \in A_{r + 1}^2, x + y \notin [\![a + b + 1, 2 a + b]\!] \implies \pi_{a, b}(x + y) \notin 
	A_{r + 1}\),
\item \(\forall c \in [\![1, c]\!], \forall x \in \TS_{b, k_c - 1}(A_c), \exists (i, j) \in {[\![1, k_c - 1]\!]}^2, i < 
	j ~\text{et}~ \pi_{a, b}(x_j - x_i) \notin A_c\).
\end{enumerate}
The set of these \(b\)-templates is denoted by \(\mathcal{T}_b(k_1, ..., k_r, t; a)\).  Color \(r + 1\) plays a special 
role and is nammed "template color".
\end{definition}

Color \(r + 1\)  is not necessarily the last color by order of appearance, the designation of \(r+1\) as template color 
symbolized by "\(t\)" is a convention that lightens notations and avoids writting 
\(\mathcal{T}_b(k_1, ..., k_{i - 1}, t,  k_{i + 1}, ..., k_{r + 1}; n)\) for instance.

If all the \(k_c\)'s are equal to 3, these \(b\)-templates can be used to produce sum-free partitions for Schur numbers.

\(b\)-Templates are a particular case of linear colorings. In the template color, the corresponding clique number is 
equal to 3. That is \(T_b(k_1, ..., k_r, t; a) \subset L(k_1, ..., k_r, 3; a + b)\).

The notion of compatibility with a \(b\)-template illustrates the fact that the final coloring in the construction is 
less constrained than the horizontal coloring, and that the constraints on this final coloring depends only on the 
horizontal coloring.

\begin{definition}[Compatibility with a \(b\)-template]
For \(A \subset \mathbb{N}\) and \(k \in \mathbb{N}\), the set of increasing sequences of length \(k\) with values in 
\(A\) is denoted by \(\mathfrak{I}_k(A)\).
